Measuring Hash Trustworthiness via Collision Utility Metrics: Logical Cryptanalysis of MD4

### 2017-12-05 dsn-2018/bare_conf.tex
Updates to future work, logical cryptanalysis techniques, and improved transition figure
Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

\subsection{Distance Metrics} \label{terminology:distance}

@@

\begin{figure*}
\begin{center}
  \begin{subfigure}{\textwidth}
    \includegraphics[width=\textwidth]{Figs/differential-1289.png}
    \caption{Collision 1289}
    \label{Fig:1289}
  \end{subfigure}\\
  \begin{subfigure}{\textwidth}
    \includegraphics[width=\textwidth]{Figs/differential-1250.png}
    \caption{Collision 1250}
    \label{Fig:1250}
  \end{subfigure}\\
  \begin{subfigure}{\textwidth}
    \includegraphics[width=\textwidth]{Figs/transitions.png}
    \caption{Differential Path Transitions}
    \label{Fig:Transitions}
  \end{subfigure}\\
\end{center}
\end{figure*}

@@

\subsection{Distance Metrics}

We find justification for the distance metric defined in Section
\ref{terminology:distance} in the existing literature on MD4:
H. Dobbertin's \cite{Dobbertin1998} and P. Kasselman's \cite{KasselmanMD4}
collision classes have distance $\delta(C_{Dobbertin}, C_{Kasselman}) = 1$
under this metric. The distance function allows us to find collision classes
based off of existing classes. By limiting ourselves to unit distances, we
maximally reuse an existing collision and limit the amount of work the SAT
solver needs to do. Further, rather than searching for any unit-distance
new collision classes, we can easily split our search to look in each of the
48 rounds of MD4, allowing us to parallelize the search for neighbors.

@@

We find justification for the distance metric $\Delta$ in the existing
literature due to collisions by  X. Wang \cite{cryptoeprint:2004:199} and
M. Schl{\"a}ffer \cite{Schlaffer2006}. Their respective families are:

\begin{align*}
F(C_{Wang}) &=& (1, 2, 3, 5, 6, 7, 8, 9, 10, 11, 12, \textbf{13}, 15, 16, 19, 20, 35, 36)\\
F(C_{Schlaffer}) &=& (1, 2, 3, 5, 6, 7, 8, 9, 10, 11, 12, 15, 16, 19, 20, 35, 36)
\end{align*}

@@

\textbf{TODO - figures and graphics on dataset}

@@

\section{Conclusion} \label{Sec:Conclusion}
The conclusion goes here.

@@

The framework can be seen here \textbf{TODO}. Complete data set available
upon request.

>>>>

\subsection{Distance Heuristics} \label{terminology:distance}

@@

\begin{figure*}
\begin{center}
  \begin{subfigure}{\textwidth}
    \includegraphics[width=\textwidth]{Figs/differential-1289.png}
    \caption{Collision 1289}
    \label{Fig:1289}
  \end{subfigure}\\
  \begin{subfigure}{\textwidth}
    \includegraphics[width=\textwidth]{Figs/differential-1250.png}
    \caption{Collision 1250}
    \label{Fig:1250}
  \end{subfigure}\\
  \begin{subfigure}{\textwidth}
    \includegraphics[width=\textwidth]{Figs/transitions.pdf}
    \caption{Differential Path Transitions}
    \label{Fig:Transitions}
  \end{subfigure}\\
\end{center}
\end{figure*}

@@

\subsection{Distance Heuristics}

We find justification for the distance heuristic, $\delta$, defined in Section
\ref{terminology:distance} in the existing literature on MD4:
H. Dobbertin's \cite{Dobbertin1998} and P. Kasselman's \cite{KasselmanMD4}
collision classes have distance $\delta(C_{Dobbertin}, C_{Kasselman}) = 1$
under this metric. Since others have shown that MD4 is untrustworthy, we can
limit our search space to small distances, exploiting the structure of MD4. This
reduction in search space decreases the time to solve a single 3-CNF-SAT
instance at the expense of restricting ourselves to highly connected, highly
similar differential paths. Further, we can achieve parallel solving
capabilities by noting that this problem is embarrassingly parallel: each
candidate collision class can be independently checked on separate machines.

The highly connected, highly similar differential paths are highlighted in
Figure \ref{Fig:Graph-Clique}: when the neighbors of H. Dobbertin's and
P. Kasselman's collision classes from Figure \ref{Fig:Graph-Intro} are viewed
independently, the large cliques become apparent. Depending on the search
strategies utilized, cliques may be expanded before new neighbors, or a
significant portion of a vertices neighbors might belong to existing cliques,
resulting in slow progress towards any external goal.

@@

We find justification for the distance heuristic $\Delta$ in the existing
literature due to collisions by  X. Wang \cite{cryptoeprint:2004:199} and
M. Schl{\"a}ffer \cite{Schlaffer2006}. Their respective families are:
$
\begin{aligned}
F(C_{Wang}) = (& 1, 2, 3, 5, 6, 7, 8, 9, 10, 11, 12, \\
        & \textbf{13}, 15, 16, 19, 20, 35, 36) \\
\end{aligned} \\
\begin{aligned}
F(C_{Schlaffer}) = (& 1, 2, 3, 5, 6, 7, 8, 9, 10, 11, 12, \\
         & 15, 16, 19, 20, 35, 36)
\end{aligned}
$

@@

\textbf{TODO - figures and graphics on dataset, cross round, etc.}

@@

\section{Conclusion} \label{Sec:Conclusion}

\textbf{TODO: COMPLETE}

@@

Moving forward, we seek to solidify the utility metrics by improving upon
second preimage and chosen preimage bounds to validate assumptions about the
utility of collisions. Further, we seek to apply these results to other hash
functions in the same family, such as MD5, SHA-1, and SHA-2. Lastly, it
remains to be seen how this work impacts dissimilar constructions such as
SHA-3 and key derivation functions such as bcrypt and scrypt, and whether
logical cryptanalysis techniques can be applied to these categories of
hash functions as well.

This research and all associated code are open access and open source.
The code is hosted on GitHub at \url{https://github.com/XXXXXXXXXXXXXXXXXX}.
For additional details or to request data associated with the project, please
contact the authors.
